### 2013-09-24 paper1.tex
intro written

\documentclass[useAMS,usenatbib]{mn2e}                                  
%
%\def\del#1{{\bf  [deleted text] }} \def\new#1{{\bf #1}} \def\remark#1{{\bf  [remark: #1]}}
%\def\del#1{} \def\new#1{#1} \def\remark#1{}

\def\del#1{{}}
%\def\del#1{{\bf (DELETED TEXT)}}
\def\C#1{{\bf #1}}
%\def\C#1{#1}

%\newcommand{\rmn}{\mathrm}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\expval}[1]{\left\langle #1 \right\rangle}
\newcommand{\RA}[3]{#1$^{\mathrm{h}}$#2$^{\mathrm{m}}$#3$^{\mathrm{s}}$}
\newcommand{\Dec}[3]{#1$^{\circ}$#2\arcmin#3\arcsec}
\newcommand{\dps}{\displaystyle}
\newcommand{\eps}{\varepsilon}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\vel}{\upsilon}
%
\usepackage{graphicx}
\usepackage{natbib}
%
\usepackage{txfonts}

%\voffset-0.5in

%MNRAS
\title[A Phenomenological Model for the Intracluster Medium]{A Phenomenological Model for the Intracluster Medium that matches X-ray and Sunyaev-Zel'dovich observations}
\author[F. Zandanel, C. Pfrommer and F. Prada]{
Fabio Zandanel$^{1,2}$, Christoph Pfrommer$^{3}$ and Francisco Prada$^{4,5,1}$\\
$^{1}$GRAPPA Institute, University of Amsterdam, Science Park 904, 1098XH Amsterdam, Netherlands, f.zandanel@uva.nl\\
$^{2}$Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (CSIC), Glorieta de la Astronom\'{\i}a, E-18080 Granada, Spain\\
$^{3}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany, christoph.pfrommer@h-its.org\\
$^{4}$Campus of International Excellence UAM+CSIC, Cantoblanco, E-28049 Madrid, Spain\\
$^{5}$Instituto de F\'{\i}sica Te\'orica, (UAM/CSIC), Universidad Aut\'onoma de Madrid, Cantoblanco, E-28049 Madrid, Spain}
\begin{document}

\date{Accepted XXX. Received XXX; in original from XXX}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2012}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf Cosmological hydrodynamical simulations of galaxy clusters are still
    challenged to produce a model for the intracluster medium that matches all
    aspects of current X-ray and Sunyaev-Zel'dovich observations. To facilitate
    such comparisons with future simulations and to enable realistic cluster
    population studies for modeling e.g., non-thermal emission processes, we
    construct a phenomenological model for the intracluster medium that is based
    on a representative sample of X-ray clusters. We create a mock galaxy
    cluster catalog by assigning our gas density model to each cluster halo of
    the large collisionless $N$-body simulation {\em MultiDark}. Our clusters
    are identified with cool-core and non cool-core clusters according to a
    dynamical disturbance parameter. We demonstrate that our gas model matches
    the various observed Sunyaev-Zel'dovich and X-ray scaling relations as well
    as the X-ray luminosity function, thus enabling a reliable mock catalog for
    present and future instruments.  In a companion paper, we apply our catalogs
    to calculate non-thermal radio and gamma-ray emission of galaxy clusters. We
    make our cosmologically complete multi-frequency mock catalogs for the
    (non-)thermal cluster emission at different redshifts publicly and freely
    available online through the MultiDark database.}
\end{abstract}

\begin{keywords}
  Galaxies: clusters: intracluster medium - X-rays: galaxies: clusters - astronomical data bases: catalogues
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}

\begingroup
\let\thefootnote\relax\footnotetext{* f.zandanel@uva.nl}
\endgroup

{\bf  Galaxy clusters are the rarest and largest gravitationally-collapsed
objects in the Universe and form at sites of constructive interference of long
waves in the primordial density fluctuations. Clusters reach radial extends of a
few Mpc and total masses $M \sim (10^{14} - 10^{15}) M_{\odot}$, of which
galaxies, hot ($1-10$~keV) gas and dark matter (DM) contribute roughly for 5, 15
and 80\%, respectively. If the thermal properties of the intracluster medium
(ICM) were solely determined by the gravity of the DM potential wells, then
cluster would obey self-similar scaling relations \citep{1986MNRAS.222..323K}.
However, X-ray observations have demonstrated that self-similarity is broken,
especially at low-mass scales (see e.g., \citealp{2005RvMP...77..207V} for a
review). Apparently the energy input from non-gravitational processes associated
with galaxy formation have a larger influence in smaller systems, i.e., at
scales of galaxy groups.

The central cooling time of the X-ray emitting gas in galaxy clusters and groups
is bimodally distributed. Approximately half of all systems have radiative
cooling times of less than 1~Gyr and establish a population of cool core (CC)
clusters while some others have cooling times that are longer than the age of
the Universe, forming the non-cool core (NCC) cluster population
\citep{2009ApJS..182...12C,2010A&A...513A..37H}. Several physical processes have
been proposed to be responsible for balancing radiative cooling of the
low-entropy gas at the centers of CCs. In particular the self-regulated nature
of the proposed feedback and observational correlations between radio activity
and central entropy (which is a proxy for a small cooling time) have put
feedback by active galactic nuclei (AGN) into the focus of recent research
\citep{2007ARA&A..45..117M, 2012NJPh...14e5023M}. The underlying physical
processes responsible for the heating include dissipation of mechanical heating
by outflows, lobes, or sound waves from the central AGN
\citep[e.g.,][]{2001ApJ...554..261C, 2002Natur.418..301B, 2002ApJ...581..223R}
or cosmic-ray heating: streaming cosmic rays excite Alfv\'en waves on which they
resonantly scatter \citep{1969ApJ...156..445K}.  Damping of these waves
transfers cosmic ray energy and momentum to the cooling plasma
\citep{1991ApJ...377..392L, 2008MNRAS.384..251G, 2011A&A...527A..99E,
  2013MNRAS.434.2209W}, heating it at a rate that balances radiative cooling on
average at each radius \citep{2013arXiv1303.5443P}.

Modeling the formation and evolution of clusters with cosmological
hydrodynamical simulations has been progressively refined over the last years to
also include (subresolution models for) AGN feedback
\citep[e.g.,][]{2007MNRAS.380..877S, 2008MNRAS.387.1403S, 2008ApJ...687L..53P,
  2012MNRAS.420.2662D, 2012MNRAS.424..190G, 2013MNRAS.428.2366V}. While the
comparison to data and simulations with AGN feedback improved for integrated
thermal properties of large cluster samples
\citep{2012ApJ...758...74B,2012ApJ...758...75B}, there are still notable
differences for differential quantities such as the pressure profile
\citep{2013A&A...550A.131P} or the profile of gas mass fractions
\citep{2012arXiv1209.4082B}. These discrepancies (that have been more dramatic
in the past without AGN feedback) motivated the development of phenomenological
models of the ICM \citep[e.g.,][]{2005ApJ...634..964O,
  2012MNRAS.422..686C}. Here, we construct a simple and purely phenomenological
model that matches all available X-ray and Sunyaev-Zel'dovich (SZ) data with the
goal to facilitate comparison of future hydrodynamical simulations, to allow the
modeling of non-thermal cluster emission over the entire electromagnetic wave
band, and finally to enable the construction of cluster mock catalogs of the
(non-)thermal cluster emission of current and future surveys. Such a model for
the ICM needs to be complementd by a sample of cluster halos, which can either
be obtained by means of analytical expressions for the cluster mass function
(e.g., \citealp{2001MNRAS.321..372J}) or cosmological cluster simulations. In
this work, we make use of the recent collisionless $N$-body simulation {\em
  MultiDark}, which employs a $\Lambda$ cold dark matter cosmology
(\citealp{2011arXiv1104.5130P}; www.multidark.org).

Our \emph{phenomenological} approach uses gas density profiles of a
representative sample of X-ray clusters, which are complemented by observed
cluster mass-dependent gas mass fractions. The resulting cluster mock sample are
then compared to data including the X-ray luminosity function (XLF), the
luminosity-mass relation, $L_{\rmn{X}}- M$, and the $Y_{\rmn{X}}-M$ relation,
where $Y_{\rmn{X}}=M_\rmn{gas}k_{\rmn{B}}T$ with an X-ray-derived gas mass
$M_{\rmn{gas}}$ and temperature $T$ \citep{2006ApJ...650..128K}. We additionally
compare our $Y_{\rmn{SZ}}-M$ relation to recent SZ-derived measurements.  In a
companion paper \citep{paper2}, we apply our ICM model to predict the
non-thermal radio halo emission a cosmological complete sample of galaxy
clusters, enabling valuable insight into the statistics of radio halos.  On of
our final data products is a complete cosmological cluster mock catalog, at
different redshifts, containing information of DM and ICM properties for each
object, together with its (non-)thermal emission at different frequencies. This
mock catalog matches existing observations and should serve eminent in making
predictions for future radio surveys.

In Section~\ref{sec:2}, we introduce the MultiDark simulation along with our
final cluster sample. We assign to each of our clusters the phenomenologically
constructed gas density profile in Section~\ref{sec:3}. In Sections~\ref{sec:4}
and \ref{sec:5}, we show that this approach can successfully reproduce the
observed X-ray and SZ cluster data. We present the resulting mock cluster
catalogs in Section~\ref{sec:6}, and eventually summarize our conclusions in
Section~\ref{sec:7}.  In this work, the cluster mass $M_{\Delta}$ and radius
$R_{\Delta}$ are defined with respect to a density that is $\Delta=200$ or
$\Delta=500$ times the \emph{critical} density of the Universe. We adopt density
parameters of $\Omega_{\rmn{m}}=0.3$, $\Omega_{\Lambda}=0.7$ and today's Hubble
constant of $H_0 = 100 \times h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} =
0.7$.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MultiDark simulation and final cluster sample}
\label{sec:2}
The MultiDark simulation used in this work is described in detail in \cite{2011arXiv1104.5130P}.  
It is an $N$-body cosmological simulation done with the Adaptive-Refinement-Tree (ART) 
code \citep{1997ApJS..111...73K} of $2048^3$ particles within a ($1000$~Mpc~$h^{-1}$)$^3$ 
cube. {\bf The adopted cosmological parameters are $\Omega_{\rmn{m}}=0.27$, $\Omega_{\Lambda}=0.73$,
$\Omega_\rmn{b}=0.0469$, $n_\rmn{s}=0.95$, $h=0.7$ and $\sigma_8=0.82$.} This simulation is 
particularly well suited for our purpose because of its large 
number of high-mass objects, i.e., galaxy clusters.

We use the MultiDark halo catalog from its on-line database,\footnote{www.multidark.org} 
constructed with the Bound Density Maxima (BDM) algorithm \citep{1997astro.ph.12217K}.  
We will mainly use $M_{500}$ and $R_{500}$ for comparison with existing observational works.  
We use the technique described in \cite{2003ApJ...584..702H} to convert $M_{200}$ and
$R_{200}$ provided by the MultiDark halo catalog to $M_{500}$ and $R_{500}$.  In
creating our cluster sample we only select distinct halos, i.e., those halos that
are not sub-halos of any other halo, which by definition are not galaxy clusters.

Additionally, we assume that the main emission mechanism in the ICM is thermal
bremsstrahlung, which is true only above a particle energy of approximately
$2.6$~keV \citep{1988xrec.book.....S}. Below this
energy, there could be other important contributions to the emission,
e.g., from atomic lines. Therefore, we impose a mass cut of
$M_{200}\geq1\times10^{14}$~$h^{-1}$~M$_{\odot}\approx1.4\times10^{14}$~$h_{70}^{-1}$~M$_{\odot}$
which ensures $k_{\rmn{B}}T \gtrsim 2.6$~keV, assuming the $M_{500} - T_{\rmn{ci}}$ relation
of \cite{2010MNRAS.406.1773M}.

{\bf In Paper~II, we present predictions for the LOFAR radio observatory
which is expected to detect diffuse radio emission in clusters up to redshift $z \approx 1$ (e.g.,
\citealp{2012JApA..tmp...34R}).} Thus, we make use of different simulation snapshots up to $z=1$. 
{\bf The extrapolation of our model beyond this redshift is rather uncertain as it is based
on observations at low(er) redshift.} In Table~\ref{tab:z}, we show the total cluster number in 
our final cluster sample at different redshifts.

\begin{table} 
\begin{center}
\caption{Number of halos in the final cluster sample}
\medskip
\begin{tabular}{cc}
\hline
\phantom{\Big|}
redshift $z$ & number of halos \\
\hline\\[-0.5em]
 0.0~~ &  13763\\
 0.1~~ &  12398\\
 0.2~~ &  10783\\ 
 0.4~~ &   ~~7789\\ 
 0.61  &  ~~5187\\ 
 0.78  &  ~~3372\\ 
 1.0~~ &  ~~1803\\[0.5em]
\hline
\end{tabular}
\label{tab:z}
\end{center}
\footnotesize{Note. We show the number of halos in our MultiDark snapshots at redshift $z$ for $M_{200}\geq1\times10^{14}$~$h^{-1}$~M$_{\odot}\approx1.4\times10^{14}$~$h_{70}^{-1}$~M$_{\odot}$. }
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Gas density modeling}
\label{sec:3}

We decided to use a \emph{phenomenological} approach and to construct the gas
density profiles directly from X-ray observations. A suitable X-ray sample that
provides the needed information is the \emph{Representative XMM-Newton Cluster
  Structure Survey} (REXCESS) sample \citep{2008A&A...487..431C,
  2009A&A...498..361P}. It is a sample of 31 galaxy clusters of different
dynamical states at redshifts $0.06<z<0.18$ with detailed information on the
de-projected electron density profiles \citep{2008A&A...487..431C}. In
Fig.~\ref{fig:gas_profiles}, we show the 31 electron density profiles of the
REXCESS sample color-coded by non cool-core clusters (NCCCs) and 
cool-core clusters (CCCs).

In order to obtain an electron density profile that we will attach to our
simulated clusters, we use a generalized Navarro-Frank-White (GNFW) profile,
\begin{equation}
n_{\rmn{e}}(x) = \frac{n_{0}}{x^{\beta}\left[1+x^{\alpha}\right]^{\frac{\delta-\beta}{\alpha}}},
\label{eq:gnfw}
\end{equation}
where $x=R/R_{\rmn{c}}$ and $R_{\rmn{c}}$ is the cluster core radius. To reduce
the dimensionality of our fit, we fix representative values of $R_{\rmn{c}} =
0.2\, R_{500}$, $\alpha = 1$ and $\delta = 2.5$. We fit the radial density
profiles of the REXCESS sample in log-log space, separating them in the two
categories of NCCCs and CCCs as shown in Fig.~\ref{fig:gas_profiles}.  We
obtain $n_{\rmn{0,NCCC}} = 1.02\times10^{-2}$~$h_{70}^{1/2}$~cm$^{-3}$,
$n_{\rmn{0,CCC}} = 8.32\times10^{-3}$~$h_{70}^{1/2}$~cm$^{-3}$,
$\beta_{\rmn{NCCC}} = -0.093$, and $\beta_{\rmn{CCC}} = 0.592$. The resulting
fits are shown in blue and red for the NCCC and CCC population, respectively.

The next step is to introduce a mass-scaling in order to apply our GNFW profiles
to all our clusters. We adopt the gas mass fraction-mass scaling,
$f_{\rmn{gas},500}-M_{500}$ of \cite{2009ApJ...693.1142S} (and adopt their equation~(8)). We can
express $f_{\rmn{gas},500}$ in the following way:
\begin{equation}
f_{\rmn{gas},500} = \frac{M_{\rmn{gas},500}}{M_{500}}  = \frac{\int_{0}^{R_{500}} \rho_{\rmn{gas}} \rmn{d}V}{M_{500}}
\label{eq:m500}
\end{equation}
with $\rho_{\rmn{gas}}(R) = \rho_{\rmn{gas}} = n_{\rmn{e}} m_{\rmn{p}} / ( X_{\rmn{H}}X_{\rmn{e}} )$ where
$m_{\rmn{p}}$ is the proton mass, $X_{\rmn{H}} = 0.76$ is the primordial hydrogen
mass fraction and $X_{\rmn{e}} = 1.157$ the ratio of electron-to-hydrogen number
densities in the fully ionized ICM \citep{1988xrec.book.....S}. For each cluster
$i$ of our sample, we then define a \emph{mass-scaled} gas profile as
$\rho_{\rmn{gas},i}=C_{i} \,\rho_{\rmn{gas}}$ with:
\begin{eqnarray}
%C_{i}  =  (0.0616\pm(0.0060g_{1}))  h_{73}^{-1.5}  \left(\frac{M_{500,i}}{10^{13} h_{73}^{-1} \rmn{M_{\odot}}}\right)^{0.135\pm(0.030g_{2})} \nonumber \\
C_{i}  & = &  (0.0656\pm(0.0064g_{1})) \, h_{70}^{-1.5}  \nonumber \\
 & \times & \left(\frac{M_{500,i}}{1.04 \times 10^{13} h_{70}^{-1} \rmn{M_{\odot}}}\right)^{0.135\pm(0.030g_{2})} \frac{M_{500,i}}{\int_{0}^{R_{500,i}} \rho_{\rmn{gas}} \rmn{d}V}
\label{eq:gas_scaling}
\end{eqnarray}
where $g_{1}$ and $g_{2}$ are random Gaussian number which we use in order to
simulate the natural scatter of the gas profiles.\footnote{The values
  $0.0064$ and $0.03$ quoted in equation~(\ref{eq:gas_scaling}) do not represent
  the proper scatter of the $f_{\rmn{gas},500}-M_{500}$ relation but reflect the
  parameter errors and we rescaled the numerical values to a Hubble constant of
  $h_{70}$ used in this work.}

\begin{figure} 
\centering
\includegraphics[width=0.48\textwidth]{figures/gas_profiles.eps}
\caption{Electron density profiles of the 31 clusters in the REXCESS sample. Grey and black lines represent 
NCCCs and CCCs, respectively. The blue and red lines represent our GNFW mean profile for the NCCCs and CCCs, respectively.}
\label{fig:gas_profiles}
\end{figure}

Hence, for each cluster in our sample we obtain a gas density profile
$\rho_{\rmn{gas},i}$ that obeys the observed $f_{gas,500}-M_{500}$ relation and
is uniquely determined by its DM mass $M_{500,i}$ and by the property of being a
NCCC or CCC. We assign the latter property to every halo depending on its
merging history. In particular, we make use of the offset parameter
$X_{\rmn{off}}$ computed for the MultiDark halo catalog. This is defined as the
distance from the halo center to the center of mass in units of the virial
radius. This parameter assesses the dynamical state of the cluster and whether
the halo experienced a recent merger or not. Current observations reveal a ratio
of NCCCs and CCCs of about $50\%$ (e.g., \citealp{2007A&A...466..805C,
  2009MNRAS.395..764S}). Since there is a correlation between merging clusters
and NCCCs, we use the median of the $X_{\rmn{off}}$ distribution to separate our
sample into CCCs and NCCCs (with NCCCs defined to be those halos with the larger
dynamical offsets). Clearly, this is an over-simplification, and future X-ray
surveys will have to determine this property also as a function of redshift.

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/lx_m.eps}
\includegraphics[width=0.48\textwidth]{figures/yx_m.eps}
%\includegraphics[width=0.33\textwidth]{figures/sz_m.eps}
\caption{{\bf X-ray scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result. The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin.  \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}.}} 
  %\emph{Right.} $Y_{\rmn{SZ}}-M_{500}$ scaling relation at $z=0.2$ in comparison
  %to the observational sample by the \protect\cite{2013arXiv1303.5080P} with a 
  %median redshift of about $0.18$. The bottom panels show the relative difference to
  %the observational scaling relations.}
\label{fig:X_LM}
\end{figure*}

We also account for redshift evolution of the gas profiles. While our NCCC and
CCC gas profiles as derived from the REXCESS cluster sample are merely used to
define a profile shape, the normalization of the gas profiles is set by the
observational $f_{\rmn{gas},500}-M_{500}$ relation \citep{2009ApJ...693.1142S}.
The 43 clusters used in \cite{2009ApJ...693.1142S} have redshifts $0.012 < z <
0.12$ with a \emph{median} of $z \approx 0.04$. Thus, our phenomenological gas
profile is representative of the cluster population at $z\approx0$. To extend this
profile to high-$z$, we include a \emph{self-similar} scaling of the gas density
as $\rho_{\rmn{gas}}(z) = E(z)^{2} \rho_{\rmn{gas}}(z=0)$, where $E(z)^{2} =
\Omega_{\rmn{m}} (1+z)^{3} + \Omega_{{\Lambda}}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{X-ray  observables}
\label{sec:4}

So far, we used a well-observed representative sample of X-ray clusters
(REXCESS) that was supplemented by X-ray data on gas mass fractions to construct
our phenomenological gas model.  Here, we will test how our cluster mock data
compares to various X-ray inferred observables. Those include the observed
$L_{\rmn{bol}} - M_{500}$ relation, the $Y_X - M_{500}$ relation, and the
XLF.

\subsection{Scaling Relations}

First, we calculate the bolometric thermal bremsstrahlung luminosity
$L_{\rmn{bol}}$ following \cite{1988xrec.book.....S}.\footnote{We check our
  procedure by fitting each of the 31 REXCESS clusters with
  equation~(\ref{eq:gnfw}) and calculating $L_{\rmn{bol}}$ with the measured gas
  temperature of each cluster. As a result, we fall short of the observed
  luminosity by a mean (median) of about $21\%$ ($20\%$). This is reasonable
  considering that we do not permit the parameters $R_{\rmn{c}}$, $\alpha$ and
  $\delta$ to vary between different objects. Additionally, we neglect atomic
  line emission which may give a noticeable contribution, in particularly for
  low-mass clusters and in the cluster outskirts of larger systems.}  To assign
a temperature to our model clusters (that is needed for calculating
$L_{\rmn{bol}}$ and $Y_{\rmn{X}}$), we adopt the $T-M_{500}$ relation by
\cite{2010MNRAS.406.1773M},
\begin{equation}
\log_{10} \left( \frac{k_{\rmn{B}}T_{\rmn{ci}}}{\rmn{keV}} \right) = 
A + B~\log_{10} \left( \frac{E(z) M_{500}}{10^{15} h_{70}^{-1} \rmn{M_{\odot}}} \right)
\label{eq:temp}
\end{equation}
where $A=0.91$, $B=0.46$, and $T_{\rmn{ci}}$ is the cluster temperature
\emph{not} centrally excised
(\citealp{2010MNRAS.406.1773M}). \cite{2010MNRAS.406.1773M} report a scatter of
$\sigma_{\rmn{yx}} = 0.06,$\footnote{Scatter is calculated as $\sigma_{\rmn{yx}}
  = \sqrt{ \left\{ \Sigma_{i=1}^{N} [Y_{i}-(A+B~X_{i})]^{2}\right\} / N-1}$
  where the sum extends over the data points $X_{i}, Y_{i}$, and $A$ and $B$ are
  the fit parameters.} which we apply to our sample using Gaussian deviates.

In the left panel of Fig.~\ref{fig:X_LM}, we show how our model
$L_{\rmn{bol}}-M_{500}$ relation compares with observations by
\cite{2010MNRAS.406.1773M} (\emph{all} data, see their Table~7). Their sample is
composed of 238 clusters at $0.02<z<0.46$ with a median of $z \approx 0.2$ and
self-consistently takes into account all selection effects, covariances,
systematic uncertainties and the cluster mass function
\citep{2010MNRAS.406.1759M}.  For this reason, we compare the
\cite{2010MNRAS.406.1773M} data to our model at $z=0.2$, and limit the
comparison to the mass range covered by the observations. Overall, there is
reassuring agreement between our phenomenological model and the data, which
probe our model most closely on scales around the cluster core radii (which is
where the contribution to $L_{\rmn{X}}$ per logarithmic interval in radius, $\dd
L_{\rmn{X}}/\dd\log r \propto r^3 n_{\rm{gas}}^2(r) \sqrt{k_{\rmn{B}}T}$,
approximately attains its maximum).  In Table~\ref{tab:LMfits}, we show our
model $L_{\rmn{bol}}-M_{500}$ scaling relation and its scatter for different
redshifts. We find that the scatter of our samples at all redshifts are Gaussian
distributed with a standard deviation of $\sigma_{yx} \approx 0.18$ that matches
the observational results of \cite{2010MNRAS.406.1773M}, which report a scatter
of $\sigma_{yx} = 0.185$.

In the right panel of Fig.~\ref{fig:X_LM}, we compare the
$Y_{\rmn{X}}-M_{500}$ relation of our sample to observational data
\citep{2010MNRAS.406.1773M}. The model agrees nicely at the high-mass end, but
underpredicts the observed scaling at low masses by about $20\%$ (at the
1-$\sigma$ level). This is the same level of deviation from the data as in the
case of $L_{\rmn{X}}$, which is more significant due to the smaller scatter in
the $Y_{\rmn{X}}$ relation. The differential contribution to the thermal energy
per logarithmic interval in radius (and hence to the integrated Compton-$y$
parameter) is given by $\dd Y /\dd\log r \propto r^3 P_{\rmn{th}}(r)$, with the
thermal gas pressure $P_{\rmn{th}}=n_{\rmn{gas}}k_{\rmn{B}}T$. It peaks at
scales slightly smaller than $R_{500}$ with 1-$\sigma$ contributions extending
out to $3\,R_{500}$ \citep{2010ApJ...725...91B}. Hence, the observational
scaling constrains our model on those large scales, quite complementary to the
X-ray luminosity. The deviations at small masses either indicates different
assumptions about $f_{\rmn{gas}}$, the gas temperature, or different selection
effects of either observational sample that we use for model calibration and
comparison.  \cite{2010MNRAS.406.1773M} determine their masses by adopting a
constant value for $f_{\rmn{gas}}$, in contrast to our approach which adopts the
observed $f_{\rmn{gas},500}-M_{500}$ relation given by
\cite{2009ApJ...693.1142S}. Additionally, we adopt the
\cite{2010MNRAS.406.1773M} \emph{centrally included} temperature throughout all
our work, while \cite{2010MNRAS.406.1773M} use the \emph{centrally excised}
temperature to calculate $Y_{\rmn{X}}$. This assumption also impacts the scatter
of the $Y_{\rmn{X}}-M$ relation. In fact, using the \emph{centrally included}
temperature, we found a scatter of $\sigma_{yx} \approx 0.11$ (see
Table~\ref{tab:YXfits} where our $Y_{\rmn{X}}$ scaling relations are reported),
significantly higher than the value of $\sigma_{yx} = 0.052$ found by
\cite{2010MNRAS.406.1773M}.

\begin{table} 
\begin{center}
\caption{$L_{\rmn{bol}}-M_{500}$ scaling relations.}
\medskip
\begin{tabular}{cccc}
\hline
\phantom{\Big|}
redshift $z$ & $A$ & $B$ & $\sigma_{yx}$ \\
\hline \\[-0.5em]
 0      & $-21.41\pm0.11$ & $1.50\pm0.01$ & 0.179\\
 0.1   & $-21.30\pm0.12$ & $1.50\pm0.01$ & 0.179\\
 0.2   & $-21.50\pm0.13$ & $1.51\pm0.01$ & 0.178\\ 
 0.4   & $-21.13\pm0.17$ & $1.49\pm0.01$ & 0.178\\ 
 0.61 & $-21.59\pm0.22$ & $1.53\pm0.01$ & 0.177\\ 
 0.78 & $-20.73\pm0.29$ & $1.48\pm0.02$ & 0.177\\ 
 1      & $-20.45\pm0.42$ & $1.46\pm0.03$ & 0.177\\[0.5em]
\hline
\end{tabular}
\label{tab:LMfits}
\end{center}
\footnotesize{Note. Scaling relations are reported in the form of $\log_{10}~(L_{\rmn{bol}}~/~E(z)~h_{70}^{-2}~10^{44}~\rmn{erg~s}^{-1})=A+B~\log_{10}~(E(z)~M_{500}~/~h_{70}^{-1}~\rmn{M_{\odot}})$. The relation scatter $\sigma_{yx}$ is also shown.}
\end{table}

\begin{table} 
\begin{center}
\caption{$Y_{\rmn{X}, 500}-M_{500}$ scaling relations.}
\medskip
\begin{tabular}{cccc}
\hline
\phantom{\Big|}
redshift $z$ & $A$ & $B$ & $\sigma_{yx}$ \\
\hline\\[-0.5em]
 0      & $-9.18\pm0.07$ & $1.61\pm0.01$ & 0.109\\
 0.1   & $-8.85\pm0.07$ & $1.59\pm0.01$ & 0.109\\
 0.2   & $-8.82\pm0.08$ & $1.59\pm0.01$ & 0.109\\ 
 0.4   & $-8.79\pm0.10$ & $1.59\pm0.01$ & 0.108\\ 
 0.61 & $-8.65\pm0.14$ & $1.59\pm0.01$ & 0.109\\ 
 0.78 & $-8.36\pm0.18$ & $1.57\pm0.01$ & 0.109\\ 
 1      & $-8.28\pm0.26$ & $1.57\pm0.02$ & 0.109\\[0.5em]  
\hline
\end{tabular}
\label{tab:YXfits}
\end{center}
\footnotesize{Note. Scaling relations are reported in the form of $\log_{10}~(E(z)~Y_{\rmn{X},500}~/~h_{70}^{-2.5}~\rmn{M_{\odot}}~\rmn{keV})=A+B~\log_{10}~(E(z)~M_{500}~/~h_{70}^{-1}~\rmn{M_{\odot}})$. The relation scatter $\sigma_{yx}$ is also shown.}
\end{table}

\subsection{Luminosity Functions}
Studies of the XLF got out of fashion during the last years due to the
difficulties of using the X-ray luminosity for cosmological purposes. The X-ray
emissivity scales with the square of the gas density, which makes it subject to
density variations and clumping.\footnote{The mean (median) difference at $z=0$
  between $L_{\rmn{bol}}$ within $R_{200}$ or within $R_{500}$ is $\approx 5\%$
  ($\approx 7\%$). While $L_{\rmn{bol}}$ refers to the quantity calculated
  within $R_{500}$, we note that the XLF for luminosities calculated within
  $R_{200}$ will be barely changed.} This implies large scatter that causes a
large Malmquist bias and underlines the necessity of careful mock surveys that
need to address all systematics.

\begin{figure} 
\centering
\includegraphics[width=0.48\textwidth]{figures/xlf.eps}
\caption{Bolometric and soft-band ($0.1-2.4$~keV) XLFs. Shown are the soft-band
  data points, the soft-band and bolometric Schechter fits of the BCS sample of
  \protect\cite{1997ApJ...479L.101E}, which has a median of $z \approx 0.08$.  While the
  soft-band XLF, which was obtained applying the \protect\cite{2010MNRAS.406.1773M}
  $L_{\rmn{X}}-M_{500}$ relation to the MultiDark $z = 0.1$ snapshot, compares
  well with the BCS data points, it deviates from the corresponding Schechter
  fit. We also show the bolometric XLF of \protect\cite{2010MNRAS.406.1773M} and
  bolometric XLF of our model at $z=0.1$. The XLFs are calculated in equally
  log-spaced mass bins; the error bars represent the Poissonian errors. Note
  that we limit the comparison to the luminosity range covered by our sample,
  where we cut the lowest part because in that range the XLF rapidly drops due
  to the imposed mass cut.}
\label{fig:XLF}
\end{figure}

Nevertheless, it provides a complementary check for our model. To this end, we
use the \emph{ROSAT} brightest cluster sample (BCS) XLF
\citep{1997ApJ...479L.101E}, which is in good agreement with results from the
\emph{ROSAT} ESO Flux-Limited X-ray (REFLEX; \citealp{2002ApJ...566...93B}) and
HIFLUGCS \citep{2002ApJ...567..716R}.  Note that the XLF is fully determined by
the mass function and the $L_{\rmn{X}}-M_{500}$ relation after taking into
account the observational biases. This means that applying the Malmquist and
Eddington-bias-corrected $L_{\rmn{X}}-M_{500}$ relation by
\cite{2010MNRAS.406.1773M} directly to the MultiDark mass function and
accounting for the observational scatter in $L_{\rmn{X}}-M_{500}$ should yield
an unbiased XLF. We show the resulting bolometric and soft-band ($0.1-2.4$~keV)
XLF in Fig.~\ref{fig:XLF} and compare those to the corresponding BCS XLFs and
to our model predictions. Note that there is only the Schechter fit available
for the BCS bolometric XLF.  While the soft-band XLF by
\cite{2010MNRAS.406.1773M} agrees well with the BCS data points, it deviates
from the corresponding Schechter fit at low luminosities. This is also true in
the bolometric band, where the XLFs of \cite{2010MNRAS.406.1773M} and our model
agree well, but deviate from the BCS Schechter fit at low luminosities. This may
be an artifact due to the use of Schechter fit instead of the data points or may
point to incompleteness of the BCS sample. Note that the Poissonian errors of
the XLF obtained from the MultiDark simulation are a lower limit as we are
neglecting the uncertainty due to cosmic variance.  Studies of the XLF will
become again an important topic with the upcoming launch of the \emph{e}ROSITA
satellite (e.g., \citealp{2011MSAIS..17..159C}) and further studies in this
direction are desirable. For these reasons, we do not show XLF predictions at
other redshifts, leaving this for a future study.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SZ scaling relations}
\label{sec:5}
In the right panel of Fig.~\ref{fig:X_LM}, we compare the
$Y_{\rmn{SZ}}-M_{500}$ relation in our model (calculated as in equation~(3) of
\citealp{2011arXiv1109.3709B}) with the observed scaling relation by the
\cite{2013arXiv1303.5080P}. {\bf We compare with the result obtained from the 
\emph{Planck} COSMO sample with Malmquist bias correction,
which has a median redshift of $0.18$. Hence, we compare the data to our relation 
at the MultiDark snapshot $z=0.2$}. Our model reproduces the data remarkably well, except 
for the high-mass end where our simulations have a weaker constraining power due to the 
smaller box size in comparison to the survey volume of {\em Planck}. We find a scatter of
$\sigma_{yx} \approx 0.11$ which compares well with the \emph{Planck} result of
$\sigma_{yx} \approx 0.08$. In Table~\ref{tab:YSZfits}, we report our SZ scaling
relations for different redshifts.

\begin{figure} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m.eps}
\caption{{\bf SZ scaling relation of our sample at $z=0.2$ in comparison to the observational 
  sample by \protect\cite{2013arXiv1303.5080P} with a median redshift of about $0.18$.
  Grey triangles show the MultiDark sample (limited to the mass range covered by observations), 
  the black line is the corresponding scaling relation, and the blue line is the observational result. 
  The black crosses represent the $Y_{\rmn{SZ}}$ median values of our sample for a given mass bin (indicated by 
  horizontal error bars), and the vertical error bars represent the standard deviation within a
  bin. The bottom panels show the relative difference to the observational scaling relations.}}
\label{fig:SZ_M}
\end{figure}

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m_SPT.eps}
\includegraphics[width=0.48\textwidth]{figures/sz_m_ACT.eps}
\caption{{\bf SZ scaling relation of our sample at $z=0.61$ compared with the SPT and ACT results. 
  Triangles show the MultiDark sample (limited to the mass range covered by observations) 
  while the crosses represent the observational result. \emph{Left.} Comparison with the 
  SPT sample \protect\citep{2013ApJ...763..127R} color-coded by redshift. 
  $Y_{\rmn{SZ, SPT}}$ is given within an aperture of  $0.2$~h$_{70}^{-1}$~Mpc.
  \emph{Right.} Comparison with the ACT sample \protect\citep{2013JCAP...07..008H} color-coded 
  by redshift. $Y_{\rmn{SZ, ACT}}$ is given within an aperture of $0.42$~h$_{70}^{-1}$~Mpc.}}
\label{fig:SZ_M_1}
\end{figure*}

\begin{table} 
\begin{center}
\caption{$Y_{\rmn{SZ}, 500}-M_{500}$ scaling relations.}
\medskip
\begin{tabular}{cccc}
\hline
\phantom{\Big|}
redshift $z$ & $A$ & $B$ & $\sigma_{yx}$ \\
\hline\\[-0.5em]
 0      & $-27.93\pm0.07$ & $1.60\pm0.01$ & 0.109\\
 0.1   & $-27.74\pm0.07$ & $1.59\pm0.01$ & 0.109\\
 0.18 & $-27.73\pm0.08$ & $1.59\pm0.01$ & 0.109\\
 0.2   & $-27.65\pm0.08$ & $1.59\pm0.01$ & 0.109\\ 
 0.4   & $-27.57\pm0.10$ & $1.59\pm0.01$ & 0.108\\ 
 0.61 & $-27.50\pm0.13$ & $1.59\pm0.01$ & 0.109\\ 
 0.78 & $-27.15\pm0.18$ & $1.58\pm0.01$ & 0.109\\ 
 1      & $-27.01\pm0.26$ & $1.58\pm0.02$ & 0.109\\[0.5em] 
\hline
\end{tabular}
\label{tab:YSZfits}
\end{center}
\footnotesize{Note. Scaling relations are reported in the form of $\log_{10}~(E(z)^{-2/3}~Y_{\rmn{SZ},500}~/~h_{70}^{-2.5}~\rmn{Mpc}^{2})=A+B~\log_{10}~(M_{500}~/~h_{70}^{-1}~\rmn{M_{\odot}})$. The relation scatter $\sigma_{yx}$ is also shown.}
\end{table}

{\bf We additionally compare our $Y_{\rmn{SZ}}$ to the results from the South Pole 
Telescope (SPT) and Atacama Cosmology Telescope (ACT) surveys.
We select a subsample of SPT clusters from \cite{2013ApJ...763..127R} by taking
only objects with confirmed redshift and signal-to-noise $\ge 5$. The resulting sample 
consists of 115 clusters with a median redshift of 0.56 and rms 0.27. Note the much larger 
scatter in redshift of the SPT sample with respect to the \emph{Planck} one that has 
a rms of 0.11. \cite{2013ApJ...763..127R} provide $Y_{\rmn{SZ}}$ 
within an aperture of $1'$, corresponding to a median radius of $0.2$~h$_{70}^{-1}$~Mpc 
at $z = 0.56$, which we use as aperture to calculate $Y_{\rmn{SZ, SPT}}$ in our MultiDark sample.

In the left panel of Fig.~\ref{fig:SZ_M_1}, we compare the SPT cluster sample to our 
scaling relation at $z = 0.61$. We color-coded the SPT clusters in order to show that the SPT 
selection function is nearly redshift-independent. The fact that the angular size of galaxy clusters 
decreases with redshift is an advantage for SPT due to its small beam, in fact, higher-redshift clusters 
are less confused by primary CMB fluctuations \citep{2013ApJ...763..127R}. This effect is inverse 
for \emph{Planck} due to its larger beam. We can conclude that while our $Y_{\rmn{SZ}}-M_{500}$ 
scaling relation compare well to the median SPT cluster population, the SPT sample has a too
large scatter to be well represented by single MultiDark snapshots.

In the right panel of Fig.~\ref{fig:SZ_M_1}, we compare our scaling relation to
the ACT sample overlapping with the Sloan Digital Sky Survey stripe 82 
\citep{2013JCAP...07..008H}. This sample has a median redshift of 0.58 with rms 
of 0.28. \cite{2013JCAP...07..008H} provides $Y_{\rmn{SZ}}$ within $R_{500}$, which
median, $0.42$~h$_{70}^{-1}$~Mpc, is almost half of ours at the
MultiDark snapshot $z=0.61$. Therefore, we use an aperture of $0.42$~h$_{70}^{-1}$~Mpc
to calculate our $Y_{\rmn{SZ, ACT}}$. 
The fact that our $R_{500}$ does not compare well with the ACT values, while it does with 
\emph{Planck}, may be explained as follows. We adopt the ACT measurements derived from 
the 148~GHz data assuming that the cluster pressure is described by the universal 
profile of \cite{2010A&A...517A..92A}. In contrast, \cite{2013arXiv1303.5080P}
derived $R_{500}$ from X-ray measurements of $M_{500}$, similarly to
our approach of constructing gas densities from an X-ray sample of clusters.

Summarizing, our SZ scaling relations compare very well to the result from
the \emph{Planck} collaboration. However, the comparison with the SPT and 
ACT samples is not as straightforward due to the different selection function 
of these telescopes with respect to \emph{Planck}, which results in a much
more (redshift) non-homogenous sample. Nevertheless, our mock clusters 
can reproduce well the median SPT and ACT populations. A more detailed
comparison would need the generation of light-cone mock catalogs,
including instrumental effects, which is beyond the scope of this 
work.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-frequency Mock Catalogs}
\label{sec:6}

{\bf As a final product of this work, we construct cosmological complete mock catalogs of galaxy clusters, at the different redshifts of 
Table~\ref{tab:z}, for $M_{200}\geq1\times10^{14}$~$h^{-1}$~M$_{\odot}\approx1.4\times10^{14}$~$h_{70}^{-1}$~M$_{\odot}$.
We make these catalogs publicly available through the MultiDark database 
(www.multidark.org) where we will also post possible updates.

Our catalogs contain all the information regarding the DM properties of each object as 
given by the corresponding original MultiDark BDM halo catalogs. Additionally, we include the
following information:

\begin{itemize}
\item $M_{500}$ and $R_{500}$ calculated from $M_{200}$ and $R_{200}$ of the BDM catalogs with the \cite{2003ApJ...584..702H} method,
\item the ICM X-ray temperature assigned via the $T_{\rmn{ci}}-M_{500}$ relation by \cite{2010MNRAS.406.1773M},
\item a state flag identifying the cluster as NCCC or CCC as described in Section~\ref{sec:3},
\item the central gas density $\rho_{\rmn{gas},0}$ with which the full ICM radial profile can be calculated as in Section~\ref{sec:3},
\item the bolometric X-ray thermal bremsstrahlung luminosity $L_{\rmn{X, bol}}$ within $R_{500}$,
\item the $Y_{\rmn{X}}$ and $Y_{\rmn{SZ}}$ parameters within $R_{500}$.
\end{itemize}

In Paper~II, we present a model to calculate the possible radio and gamma-ray emission of 
clusters as result of the cosmic-ray (CR) proton-proton interactions with the ICM. We refer
to that paper for all the details. However, we want to point out that our mock catalogs include also:

\begin{itemize}
\item a state flag identifying whether a given object is a radio-loud or a radio-quiet cluster, 
\item the CR-to-thermal pressure averaged over the cluster volume within $R_{500}$,
\item the synchrotron radio luminosity, within $R_{500}$, at 120~MHz and 1.4~GHz,
\item the gamma-ray luminosity, within $R_{500}$, above 100~MeV and 100~GeV.
\end{itemize}

We warn, however, that our mock catalogs are not appropriate for all purposes. In fact, they are 
representations of the average cluster population. While they are very useful as baseline for new 
simulations, to test new observations, and for the non-thermal modeling of clusters, they are not 
appropriate to perform detailed cosmology studies. A carefully modeling of the response function 
of the considered instrument would be needed for such a purpose, together with the generation of 
light-cone mock catalogs. Nevertheless, for flux-complete surveys, our mock catalogs are
certainly good tools to test against models. Finally, keep also in mind that our approach is based on 
X-ray observations of clusters at low redshift \citep{2008A&A...487..431C,2009ApJ...693.1142S}, 
and, for simplicity, we adopted a self-similar redshift evolution.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:7}

{\bf We build a complete cosmological sample of galaxy clusters from the MultiDark 
$N$-body simulation with redshifts ranging from $z = 0$ to 1. We construct a 
\emph{phenomenological} model for the cluster-gas distribution. This is characterized 
by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample) 
and a cluster-mass-dependent gas fraction. We assign such a (cluster mass-dependent) 
gas density profile to each DM halo in our sample and sort it into the NCCC/CCC populations 
according to a dynamical disturbance parameter that is calculated from the DM distribution.
We obtain cosmologically complete mock catalogs of galaxy clusters that matches the 
observed $L_{\rmn{X, bol}}$-to-$M_{500}$, $Y_{\rmn{X}}$-to-$M_{500}$, and 
$Y_{\rmn{SZ}}$-to-$M_{500}$ relations, as well as the X-ray luminosity function. 
In the companion Paper~II, we additionally calculate the possible radio and gamma-ray 
emission coming from the CR hadronic interactions with the ICM in the clusters.

Eventually, we provide a cosmologically complete multi-frequency mock catalog for the (non-)thermal 
cluster emission at different redshifts. We make these catalogs publicly and freely available 
on-line through the MultiDark database (www.multidark.org). We warn that these catalogs
are not appropriate for detailed cosmological studies in their (present) published form as 
carefully modeling of the response function of the considered instrument would be needed 
for such a purpose, together with the generation of light-cone mock catalogs. Note also that 
our approach is based on X-ray observations at low redshift, and we adopted a self-similar redshift 
evolution. Nevertheless, our mock catalogs are good representations of the average cluster population,
and we hope that the  community can make valuable use of these catalogs in synergy with the future radio, 
X-ray and gamma-ray data.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
We thank Harald Ebeling, Stefan Gottl{\"o}berg, Anatoly Klypin, Andrey Kravtsov, Adam Mantz, 
Frazer Pearce, Jos\'e Alberto Rubi\~no, and Gustavo Yepes, for the useful discussions and advices. 
We also thank the MultiDark database people, in particular Adrian Partl and Kristin Riebe. 
F.Z.{\ }acknowledges the CSIC financial support as a JAE-Predoc grant of the 
program ``Junta para la Ampliaci\'on de Estudios'' co-financed by the FSE. 
F.Z.{\ }acknowledges the hospitality of the Leiden Observatory during his stay.
F.Z.{\ }and F.P.{\ } thank the support of the Spanish MICINN's Consolider-Ingenio 2010 Programme under
grant MultiDark CSD2009-00064. C.P.{\ }gratefully acknowledges financial support
of the Klaus Tschira Foundation. The MultiDark Database used in this paper and
the web application providing online access to it were constructed as part of
the activities of the German Astrophysical Virtual Observatory as result of a
collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and
the Spanish MultiDark Consolider Project CSD2009-00064. The Bolshoi and
MultiDark simulations were run on the NASA's Pleiades supercomputer at the NASA
Ames Research Center.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{mn2e}
\bibliography{bib_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{lastpage}

\end{document}

\del{
The study of the (non-)thermal emission in galaxy clusters provides
important information on structure formation process and on 
fundamental characteristics of our Universe. This field is at a 
crucial point of its development thanks to the very low sensitivity of present and future-generation 
instruments such as LOFAR\footnote{www.lofar.org}, \emph{e}ROSITA\footnote{www.mpe.mpg.de/eROSITA}, 
ASTRO-H\footnote{astro-h.isas.jaxa.jp} and CTA\footnote{www.cta-observatory.org}, to name but a few. 
Therefore, it is important to provide prospects for future surveys and
to understand in which way they can improve our current knowledge.
To this end, a complete cosmological mock sample of galaxy
clusters is needed. This can be realized either using analytical 
expressions for the cluster mass function (e.g., \citealp{2001MNRAS.321..372J}) or cosmological cluster 
simulations. In this work, we make use of the recent MultiDark $N$-body
simulation (\citealp{2011arXiv1104.5130P}; www.multidark.org).
%
Our main motivation to build such a mock catalog has been to predict the 
non-thermal emission at radio frequencies of a cosmological 
complete sample of galaxy clusters. This is presented in a companion 
paper (\citealp{paper2}; hereafter Paper~II). As described in Paper~II, for 
any given cluster, fundamental ingredients for the non-thermal emission 
model are the mass and gas density distribution. The MultiDark 
simulation used here is a $N$-body, \emph{dark matter (DM)-only} simulation and, 
therefore, does not contain any information on the intra-cluster medium (ICM) gas  
distribution. A cosmological complete sample of clusters can 
only be obtained from $N$-body simulations as hydrodynamical
simulations of such large volumes are computationally too expensive.
Therefore, we need a method to determine the ICM properties
of any given cluster from its total mass. 
}